---
title: "Analyse compatible motif-domain interactions from peptide array results"
output:
  html_document:
    code_folding: show
    style: spacelab
    toc: yes
    toc_float: yes
params:
  workdir: '.'
  datadir: './data'
date: "`r date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 12, fig.height = 10, eval = TRUE)
#knitr::opts_knit$set(root.dir = '/data/akalin/buyar/collaborations/selbach/arrayResultsAnalysis/NeuroarrayResults_20170509/')
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

```{r input_processing, eval = TRUE}
library(data.table)
library(DT)
library(Biostrings)
library(ggplot2)
library(pheatmap)
library(parallel)
library(ggsignif)
library(igraph)
library(cowplot)
library(ggnetwork)
library(intergraph)

datadir <- '/data/local/buyar/collaborations/katrina/collab_katrina_selbach_glut1/data'

#1. read peptide array interaction data
arrayData <- fread(file.path(datadir, '20170522_Neuroarray_results.preprocessed.tsv'))
arrayData$Genename.prey <- gsub(';.*$', '', arrayData$Genename.prey)

#get slim domain interactions (output of findSLiMDomainPairs.R)
slimDomainInteractions <- readRDS(file.path(datadir, 'peptideArray_slimDomainInteractions.RDS'))
slimDomainInteractions$PeptideIDext <- arrayData[match(slimDomainInteractions$interactionID, arrayData$interactionID)]$PeptideIDext
```

`r date()`

# Proteins targeting multiple peptides

We'd like to summarize for each protein the list of peptides with significant interactions with significant SILAC values

```{r}
#subset arrayData for significant interactions

mdata <- arrayData[!is.na(`Median.SILAC.ratio.Wt/Mut`)]

mdata <- rbind(arrayData[genotype == 'wt' & LFQ_strict == '+' & 
                           `Median.SILAC.ratio.Wt/Mut` > 1 & 
                           `Minimum.SILAC.ratio.Wt/Mut` > 0], 
               arrayData[genotype == 'mut' & LFQ_strict == '+' & 
                           `Median.SILAC.ratio.Wt/Mut` < -1  & 
                           `Maximum.SILAC.ratio.Wt/Mut` < 0])

#split mdata by prey gene name
ldata <- lapply(split(mdata, mdata$Genename.prey), function(x) unique(x$PeptideIDext))
ldata <- ldata[names(sort(sapply(ldata, length), decreasing = T))]

dt <- data.table::data.table(do.call(rbind, lapply(names(ldata), 
                      function(x) {
                        data.frame('geneName' = x,
                                   'peptides' = paste0(ldata[[x]], collapse = ', '),
                                   'peptideCount' = length(ldata[[x]]),
                                   stringsAsFactors = F)
                         })))

DT::datatable(data = dt,
          extensions = c('Buttons', 'FixedColumns'),
          options = list(fixedColumns = TRUE,
                         scrollX = TRUE,
                         dom = 'Bfrtipl',
                         buttons = c('copy', 'print', 'csv','excel', 'pdf')),
          filter = 'bottom')
```


# How specific are the pull-downs? 

We'd like to find out how specific are the interactions detected from the array.
Are the interactions that we call significant, enriched for relevant protein
domains that can bind slims whose patterns match the peptides?

## Does gain/loss of slim-domain interactions correlate with the changes in the silac ratios? 

Assuming that the availability of a SLiM in a peptide and a compatible PFAM
domain in its interaction partner provides us evidence of the accuracy of the
detected interaction, then we should observe an impact in the binding strength
between the peptide and its partner when the SLiMs are broken by the mutations,
or novel slims are gained via mutations.

In the following analysis, I look for interactions of peptides in which a
SLiM-domain pair is broken (lost interaction for the mutant peptide) and
interactions that have created a SLiM-domain pair (gained interaction for the
mutant peptide) and check if the SILAC ratios (wt versus mutant) have
decreased/increased in significant interactions and compare it with the
background distribution of SILAC ratios.

SILAC values are plotted for:

- all-interactions: all detected interactions for WT/mutant peptides

- slim-domain-paired: interactions in which the peptide has a SLiM and the
interaction partner has a compatible PFAM domain

- slim-domain-pair_GAINED: interactions in which the mutant peptide has gained a
SLiM (via point mutation) and the interaction partner contains a compatible PFAM
domain

- slim-domain-pair_LOST: interactions in which the mutant peptide has lost a
SLiM (via point mutation) and the interaction partner contains a compatible PFAM
domain


```{r}
all <- arrayData[!is.na(`Median.SILAC.ratio.Wt/Mut`)] #all interactions
all$category <- 'All interactions'

gained <- arrayData[genotype == 'mut' & !is.na(`Median.SILAC.ratio.Wt/Mut`) & interactionID %in% slimDomainInteractions[gainedTargets != '']$interactionID]

gainedN <- gained[LFQ_loose == '-']
gainedN$category <- 'Gained interactions: LFQ negative'

gainedP <- gained[LFQ_loose == '+']
gainedP$category <- 'Gained interactions: LFQ positive'


lost <- arrayData[genotype == 'wt' & !is.na(`Median.SILAC.ratio.Wt/Mut`) & interactionID %in% slimDomainInteractions[lostTargets != '']$interactionID]

lostN <- lost[LFQ_loose == '-']
lostN$category <- 'Lost interactions: LFQ negative'

lostP <- lost[LFQ_loose == '+']
lostP$category <- 'Lost interactions: LFQ positive'

mdata <- rbind(all, gainedN, gainedP, lostN, lostP)
counts <- mdata[,length(PeptideID), by = c('category')]

ggplot(data = mdata, aes(x = category, LFQ_loose, y = `Median.SILAC.ratio.Wt/Mut`)) + 
  theme_bw(base_size = 16) +
  geom_violin(aes(fill = category)) + 
    geom_signif(comparisons = list(
            c('Gained interactions: LFQ negative', 'All interactions'),
            c('Gained interactions: LFQ positive', 'All interactions'),
            c('Lost interactions: LFQ negative', 'All interactions'),
            c('Lost interactions: LFQ positive', 'All interactions')),
      map_signif_level = FALSE, y_position = c(8:11)) + 
    geom_text(data = counts, aes(x = category, y = -8, label = paste0("N: ",V1))) +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))
```

# Gained/Lost SLiM-domain interactions 

## SLiM-domain interactions lost in the mutant
```{r gained-lost_slimdomain_interacitons}

slimDomainInteractions$`Median.SILAC.ratio.Wt/Mut` <- arrayData[match(slimDomainInteractions$interactionID, arrayData$interactionID)]$`Median.SILAC.ratio.Wt/Mut`
slimDomainInteractions$LFQ_loose <- arrayData[match(slimDomainInteractions$interactionID, arrayData$interactionID)]$LFQ_loose
slimDomainInteractions$genotype <- arrayData[match(slimDomainInteractions$interactionID, arrayData$interactionID)]$genotype
slimDomainInteractions$LFQ_strict <- arrayData[match(slimDomainInteractions$interactionID, arrayData$interactionID)]$LFQ_strict


lostInteractions <- slimDomainInteractions[lostTargets != '' & LFQ_loose == '+' & !is.na(`Median.SILAC.ratio.Wt/Mut`) & genotype == 'wt'] 

lostInteractions <- subset(lostInteractions, select = c('PeptideID', 'domainName', 'domainTargetsInPeptideWT', 'domainTargetsInPeptideMUT', 'interactor', 'Median.SILAC.ratio.Wt/Mut'))

DT::datatable(data = lostInteractions,
          extensions = c('Buttons', 'FixedColumns'),
          options = list(fixedColumns = TRUE,
                         scrollX = TRUE,
                         dom = 'Bfrtipl',
                         buttons = c('copy', 'print', 'csv','excel', 'pdf')),
          filter = 'bottom'
          )
```

## SLiM-Domain interactions gained in the mutant

```{r}
gainedInteractions <- slimDomainInteractions[gainedTargets != '' & LFQ_loose == '+' & !is.na(`Median.SILAC.ratio.Wt/Mut`) & genotype == 'mut'] 

gainedInteractions <- subset(gainedInteractions, select = c('PeptideID', 'domainName', 'domainTargetsInPeptideWT', 'domainTargetsInPeptideMUT', 'interactor', 'Median.SILAC.ratio.Wt/Mut'))

DT::datatable(data = gainedInteractions,
          extensions = c('Buttons', 'FixedColumns'),
          options = list(fixedColumns = TRUE,
                         scrollX = TRUE,
                         dom = 'Bfrtipl',
                         buttons = c('copy', 'print', 'csv','excel', 'pdf')),
          filter = 'bottom'
          )
```

# Clustering the peptide-protein network

Here, I generate a network of peptide-protein interactions using LFQ_strict cut-off and SILAC thresholds. 
Then, I split the network into communities using the igraph package and find the communities with at least two peptides. 
Then, I look for common matches of SLiMs in these clustered peptides and for each cluster I print the top 3 SLiMs sorted by number of peptides that they match. 

```{r clusterNetwork, eval = TRUE}

minMembers <- 1 

mdata <- rbind(arrayData[genotype == 'wt' & LFQ_strict == '+' & 
                           `Median.SILAC.ratio.Wt/Mut` > 1 & 
                           `Minimum.SILAC.ratio.Wt/Mut` > 0], 
               arrayData[genotype == 'mut' & LFQ_strict == '+' & 
                           `Median.SILAC.ratio.Wt/Mut` < -1  & 
                           `Maximum.SILAC.ratio.Wt/Mut` < 0])

#exclude control peptides from network analysis
mdata <- mdata[!grep('cntrl', mdata$PeptideID),]

mdata <- subset(mdata, select = c('Genename.prey', 'PeptideIDwithGeneName', 'genotype', 'Median.SILAC.ratio.Wt/Mut', 'uniprotMatch', 'PeptideUniprotID'))
colnames(mdata)[1:2] <- c('from', 'to')

#create a network and get communities
g <- graph_from_data_frame(d = mdata, directed = F)
fc <- fastgreedy.community(as.undirected(g))
#get community membership of each node into a data.frame
memb <- data.frame(t(data.frame(as.list(membership(fc)))))
colnames(memb) <- c('cluster')
memb$nodes <- rownames(memb)

#define a mapping between abbreviated source names from gprofiler to more meaningful names
domainNames <- list('BP' = 'GO:BP', 
                    'MF' = 'GO:MF', 
                    'CC' = 'GO:CC', 
                    'keg' = 'KEGG', 
                    'rea' = 'REACTOME')

# run gprofiler for each subcluster for GO terms, KEGG and REACTOME pathways. 
goResults <- pbapply::pbsapply(simplify = FALSE, USE.NAMES = TRUE, 
                                X = split(memb, memb$cluster), function(x) {
  #remove peptides from vertices to only consider interaction partners of peptides
  vertices <-  setdiff(x$nodes, mdata$to)
  #look for go terms when there is at least two genes in the subgraph
  if(length(vertices) > 1) {
    ids <- unique(mdata[from %in% vertices]$uniprotMatch)
    res <- gProfileR::gprofiler(query = ids, organism = 'hsapiens', src_filter = c('GO', 'KEGG', 'REAC'))
    res$domain <- as.vector(unlist(domainNames[res$domain]))
    if(nrow(res) > 0) {
      return(res)
    } else {
      return(NULL)
    }
  } else {
    return(NULL)
  }
})

#for each subgraph plot the interaction network  
subgraphs <- lapply(clusterIDs, function(clusterID) {
#plot a subgraph of g based on membershipd data 
cat("clusterID",clusterID,"\n")
vertices <- memb[memb$cluster == clusterID,]$nodes
if(length(vertices) > 2) {
  d <- mdata[to %in% vertices]
  gi <- graph_from_data_frame(d = d, 
                              directed = F)
  igraph::V(gi)$color <- ifelse(grepl('_', names(V(gi))), "peptide", "protein")
  igraph::V(gi)$shape <- ifelse(grepl('_', names(V(gi))), "peptide", "protein")
  igraph::E(gi)$color <- ifelse(d$genotype == 'wt', "wt-edge", "mutant-edge")
  igraph::E(gi)$width <- abs(d$`Median.SILAC.ratio.Wt/Mut`)
  gi <- add_layout_(gi, nicely())
  
  p <- ggplot(ggnetwork(gi), aes(x = x, y = y, xend = xend, yend = yend), show.legend = F) +
      geom_edges(aes(color = color.y)) +
      geom_nodes(aes(color = color.x, 
                     shape = ifelse(shape == 'protein', 'circle', 'triangle')), 
                 show.legend = F,
                 size = 8) +
      geom_nodelabel_repel(aes(color = color.x, label = vertex.names), 
                           show.legend = F, size = 3,
                           fontface = "bold") +
      theme_blank(legend.position = 'bottom', 
                  legend.box.margin = margin(0, 0, 0, 0), 
                  legend.title = element_blank())
  return(p)
  }
})
names(subgraphs) <- clusterIDs

# make plots for each sub graph 
pdf("network_data.clustering_goterms.pdf")
i <- 0
for(clusterID in names(goResults)) {
  if(!is.null(subgraphs[[clusterID]]) & !is.null(goResults[[clusterID]])) {
    i <- i + 1
    sg <- subgraphs[[clusterID]]
    goData <- goResults[[clusterID]]
    goData$term.name <- stringr::str_wrap(factor(goData$term.name), width = 30)
    if(nrow(goData) > 10) { goData <- goData[1:10,] } 
    go <- ggplot(data = goData, aes(x = term.name, y = -log10(p.value))) +
      geom_bar(aes(fill = domain, reorder(term.name, p.value)), stat = 'identity', width = 0.25) + 
      theme(axis.text.x = element_text(angle = 45, size = 8, hjust = 0.8, vjust = 0.9),
            axis.title.x = element_blank(), axis.title.y = element_text(size = 8), 
            legend.title = element_text(size = 8), legend.text = element_text(size = 8), 
            legend.position = c(0.9,0.9), aspect.ratio = 0.25)
    print(cowplot::plot_grid(sg, go, nrow = 2) + draw_label(label = LETTERS[i],
                                                      x = 0.05, y = 0.95))
  }
}
dev.off()
```
